---
title: "Gamification Analysis"
author: "Alexander (Sasha) Pastukhov"
date: "11 Juli 2018"
output:
  github_document:
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    keep_md: true
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library('tidyverse')
library('knitr')
library('lme4')
library('lmerTest')
library('BayesFactor')
library('psycho')
library('coin')
library('car')
library('lmPerm')
library('boot')
```

# Loading and preprocessing the data

```{r Loading}
rm(list= ls())

results <- tibble(filename= list.files(path= 'Data', pattern = 'csv$')) %>%
  group_by(filename) %>%
  do(suppressMessages(read_csv2(file.path('Data', .)))) %>%

  # converting selected variables to factors and using meaningful labels instead of boolean balues
  ungroup() %>%
  mutate(COA= as.factor(COA), 
        Target= as.factor(Target), 
        Condition= as.factor(Condition),
        Observer= as.factor(Observer),
        
        Match= ifelse(Match == 'True', 'match', 'mismatch'), 
        Response= ifelse(Response == 'True', 'match', 'mismatch'), 
        Correct= Match == Response)

# adding block index within the COA
results <- results %>%
  group_by(Observer, Condition, COA, Block) %>%
  summarise() %>%
  
  group_by(Observer, Condition, COA) %>%
  mutate(coaBlockIndex= 1:n()) %>%

  right_join(results, by= c('Observer', 'Condition', 'COA', 'Block')) %>%

    # retain only 12 blocks to facilitate comparison (some observers had more blocks)
  filter(Block <= 12)
```

Computing score for each trial, including the _experiment_ condition group. Later, this would show whether having a maximal score in mind alters your behavior. For each trial the score is calculated as $$S(t)= Combo(t) \times \frac{ 10 \times (4-RT(t))}{3}$$/, where _RT(t)_ is the response time and _Combo(t)_ is a combo multiplier. The latter is increased by 1 after every correct answer but is reset to 1 after a mistake. 

```{r Compute score}

# computing time-based score, maximum is 10, but time penalty applies
# this value will be translated into the real score later, based on combo and response
results <- results %>%
  # computing the time-based score the way it was done in the game
  mutate(timebased_score= ceiling(10*((4-RT)/3))) %>%
         
  # clamping the score within 0..10 range, -1 means that the response was incorrect
  mutate(timebased_score= ifelse(timebased_score>10, 10, timebased_score),
         timebased_score= ifelse(timebased_score<0, 0, timebased_score), 
         timebased_score= ifelse(!Correct, -1, timebased_score))
         

# computing combo-based score         
compute_score_within_block <- function(timebased_score){
  combo <- 1
  score <- rep(NA, length(timebased_score))
  for(iTrial in 1:length(timebased_score)){
    if (timebased_score[iTrial]>=0){
      score[iTrial] <- timebased_score[iTrial] * combo
      combo <- combo + 1
    }
    else{
      score[iTrial] <- 0
      combo <- 1
    }
  }
  
  score
}


results <- results %>%
  group_by(Observer, Condition, COA, Block) %>%
  mutate(score= compute_score_within_block(timebased_score))

rm('compute_score_within_block')
```


# Comparing two experimental groups to see if having add-on gamification changes behaviors

## Effect of the cue-onset asynchrony (COA) and experimental condition on _performance_
```{r Function to bootstrap CI for group means}
#' Computes mean value for the sample
#'
#' @description Computes mean value for the \code{variableOfInterest} using a sample of row, 
#'  defined by \code{bootstrapIndex}, from the \code{variableOfInterest}. 
#' @param irrelevantData fake data supplied to \code{boot} function those length is equal to the total 
#' number of rows for a single observer 
#' @param bootstrapIndex index of rows supplied by \code{boot} function.
#' @param dfOfInterest table with actual group data
#' @param variableOfInterest name of the column that must be sampled
#' @param avgFun function used for computing average value
#'
#' @return sampled group mean 
getVariableSampleMean <- function(irrelevantData, bootstrapIndex, dfOfInterest, variableOfInterest, avgFun){
  dfOfInterest %>%
    # first sampling each observer the same way and computing their averages
    group_by(Observer) %>%
    slice(bootstrapIndex) %>%
    summarise(varMean= avgFun(!!as.name(variableOfInterest), na.rm= TRUE)) %>%
  
    # then, the group average
    ungroup() %>%
    summarise(varMean= mean(varMean)) %>%
    
    # returning only the group mean
    pull(varMean)
}

#' Computes variable mean and 95% bootstrapped bca CI 
#'
#' @param data data table
#' @param variableOfInterest name of the column to analyze
#' @param seed seed for the random numbers generator (to make CIs reproducable). NULL means no seeding
#' @param avgFun function used for computing average value, defaults to 'mean'
#' @param R number of bootstrap iterations, defaults to 2000
#'
#' @return table with columns bcaLower, meanValue, bcaUpper
getvariableCI <- function(data, variableOfInterest, seed= NULL, avgFun= mean, R= 2000){
  # figuring out number of trials per observer
  trialN <- data %>%
    group_by(Observer) %>%
    summarise(trialCount= n()) %>%
    pull(trialCount)
  
  # seeding the random numbers generator
  if (!is.null(seed)){
    set.seed(seed)
  }
  
  # sample mean
  sampledMean <-boot(data= 1:max(trialN), 
                     statistic = getVariableSampleMean, R= R, 
                     dfOfInterest= data, 
                     variableOfInterest= variableOfInterest, 
                     avgFun= avgFun)
  
  # compute CIs
  sampledCI <- boot.ci(boot.out= sampledMean, type= 'bca')
  
  # compute mean over ALL the data
  meanValue <- getVariableSampleMean(NULL, 1:max(trialN), data, variableOfInterest, avgFun)
  
  # package for output
  data.frame(bcaLower= sampledCI$bca[4], meanValue= meanValue, bcaUpper= sampledCI$bca[5])
}
```


```{r, Group Performance, cache=TRUE}
groupPerformance <- results %>%
  group_by(Condition, COA, coaBlockIndex) %>%
  do(getvariableCI(data= ., variableOfInterest = 'Correct', seed = 1538985527))

ggplot(data= groupPerformance, 
       aes(x= coaBlockIndex, y= meanValue, ymin= bcaLower, ymax= bcaUpper,
           color= Condition, linetype= Condition)) + 
  geom_errorbar(width= 0.3, linetype= 'solid')+
  geom_line() +
  geom_point(aes(shape= Condition), size= 3) +
  # geom_point(position = position_jitter(width= 0.2), color= 'black') +
  facet_grid(.~COA) +
  ylab('Performance [%]') +
  xlab('Block index') +
  theme(panel.grid.minor.x = element_blank(), legend.position = "none")
```

__Performance: Frequentist ANOVA__
```{r, Performance - frequentist ANOVA}
performance_coa <- results %>%
  group_by(Observer, Condition, COA, coaBlockIndex) %>%
  summarise(Performance= 100 * mean(Correct), 
            logitPerformance= car::logit(Performance, percents = TRUE, adjust = 0.025))

# frequentist ANOVA
kable(summary(aov(logitPerformance ~ coaBlockIndex * COA * Condition, data= performance_coa))[[1]])
```


**Performance: Permutation ANOVA**
```{r, Performance - permutation ANOVA}
kable(summary(aovp(Performance ~ coaBlockIndex * COA * Condition, data= performance_coa))[[1]])
```

**Performance: Bayesian ANOVA**

```{r, Performance - bayesian ANOVA}
performance_coa %>%
  ungroup() %>%
  mutate(coaBlockIndex= as.factor(coaBlockIndex)) %>%
  anovaBF(logitPerformance ~ coaBlockIndex * COA * Condition, data= .) %>% 
  extractBF(.) %>%
  select(-time, -code) %>%
  kable(.)
```

```{r, Performance - clean up}
rm(groupPerformance, performance_coa)
```

## Effect of the cue-onset asynchrony (COA) and experimental condition on _response times_

```{r, group RT, cache= TRUE}
groupMedianRT <- results %>%
  group_by(Condition, COA, coaBlockIndex) %>%
  do(getvariableCI(data= ., variableOfInterest = 'RT', seed = 1538988495, avgFun = median))


ggplot(data= groupMedianRT, 
       aes(x= coaBlockIndex, y= meanValue, ymin= bcaLower, ymax= bcaUpper,
           color= Condition, linetype= Condition)) + 
  geom_errorbar(width= 0.3, linetype= 'solid')+
  geom_line() +
  geom_point(aes(shape= Condition), size= 3) +
  # geom_point(position = position_jitter(width= 0.2), color= 'black') +
  facet_grid(.~COA) +
  ylab('Median RT [s]') +
  xlab('Block index') +
  theme(panel.grid.minor.x = element_blank(), legend.position = "none")


ggplot(data= groupMedianRT, 
      aes(x= coaBlockIndex, y= RT_avg, ymin= RT_lo, ymax= RT_hi,
         color= Condition, linetype= Condition)) + 
  geom_errorbar(width= 0.3)+
  geom_line() +
  geom_point(aes(shape= Condition), size= 3) +
  # geom_point(position = position_jitter(width= 0.2), color= 'black') +
  facet_grid(.~COA) +
  ylab('Median RT [s]') +
  xlab('Block index') +
  theme(panel.grid.minor.x = element_blank(), legend.position = "none")

print(RT_plot)

# ggsave('Between-RT.pdf', RT_plot, path= 'Plots', width = 20, height = 20/1.5, units = 'cm', useDingbats= FALSE)
```

__RT: Frequentist ANOVA__
```{r, RT - frequentist ANOVA}
# frequentist ANOVA

RT_coa <- results %>%
  group_by(Observer, Condition, COA, coaBlockIndex) %>%
  summarise(RT= median(RT))

kable(summary(aov(RT ~ coaBlockIndex * COA * Condition, data= RT_coa))[[1]])
```


**RT: Permutation ANOVA**
```{r COA on RT - permutation ANOVA}
kable(summary(aovp(RT ~ coaBlockIndex * COA * Condition, data= RT_coa))[[1]])
```

**RT: Bayesian ANOVA**

```{r COA on RT - bayesian ANOVA}
RT_coa %>%
  ungroup() %>%
  mutate(coaBlockIndex= as.factor(coaBlockIndex)) %>%
  anovaBF(RT ~ coaBlockIndex * COA * Condition, data= .) %>% 
  extractBF(.) %>%
  select(-time, -code) %>%
  kable(.)
```

```{r COA on RT - clean up}
rm(RT_coa, RT_plot)
```

## Effect of the cue-onset asynchrony (COA) and experimental condition on _total score_
```{r COA on Score}
score_coa <- results %>%
  group_by(Observer, Condition, COA, coaBlockIndex) %>%
  summarise(`Block Score`= sum(score))

# plot with SE as errorbars
score_plot <- score_coa %>%
  group_by(Condition, COA, coaBlockIndex) %>%
  summarise(Score_avg= mean(`Block Score`), 
            Score_serr= sd(`Block Score`)/sqrt(n()-1), 
            Score_lo= Score_avg-Score_serr, 
            Score_hi= Score_avg+Score_serr) %>%

ggplot(aes(x= coaBlockIndex, y= Score_avg, ymin= Score_lo, ymax= Score_hi,
           color= Condition, linetype= Condition)) + 
  geom_errorbar(width= 0.3)+
  geom_line() +
  geom_point(aes(shape= Condition), size= 3) +
  # geom_point(position = position_jitter(width= 0.2), color= 'black') +
  facet_grid(.~COA) +
  ylab('Score [points]') +
  xlab('Block index') + 
  theme(panel.grid.minor.x = element_blank(), legend.position = "none")
print(score_plot)
# ggsave('Between-Score.pdf', score_plot, path= 'Plots', width = 20, height = 20/1.5, units = 'cm', useDingbats= FALSE)

```


**Score: Frequentist ANOVA**
```{r COA on Score - frequentist ANOVA}
# frequentist ANOVA
# frequentist ANOVA
kable(summary(aov(`Block Score` ~ coaBlockIndex * COA * Condition, data= score_coa))[[1]])
```


**Score: Permutation ANOVA**
```{r COA on Score - permutation ANOVA}
kable(summary(aovp(`Block Score` ~ coaBlockIndex * COA * Condition, data= score_coa))[[1]])
```

**Score: Bayesian ANOVA**

```{r COA on Score - bayesian ANOVA}
score_coa %>%
  ungroup() %>%
  mutate(coaBlockIndex= as.factor(coaBlockIndex)) %>%
  anovaBF(`Block Score` ~ coaBlockIndex * COA * Condition, data= .) %>% 
  extractBF(.) %>%
  select(-time, -code) %>%
  kable(.)
```
```{r COA on Score - cleanup}
rm(score_coa, score_plot)
```



# Post-study questionnaires

``` {r Load and preprocess questionnaires}
questionnaire <- suppressMessages(read_csv2(file.path('Questionnaire', 'Questionnaire_ErsteTestung_ohneIFN.csv'))) %>%
  ungroup() %>%
  mutate(Condition= as.factor(Condition), 
         Condition= fct_relevel(Condition, 'game'))
```

## Gamer/Non-gamer

This questionnaire is important to check the balance between the two experimental groups

```{r Gamer}
gamer <- questionnaire %>%
  mutate(d_nonGamer= sqrt((GP01_01-2.58)^2+(GP02_01-1.61)^2+
                     (GP02_02-1.88)^2+(GP02_03-4.3)^2+
                     (GP02_04-1.49)^2+(GP02_05-2.69)^2+
                     (GP02_06-2.20)^2+(GP02_07-2.59)^2+
                     (GP02_08-4.68)^2+(GP02_09-2.20)^2)) %>%
  select(Observer, Condition, d_nonGamer)

oneway_test(d_nonGamer ~ Condition, gamer)


gamer_plot <- gamer %>% ggplot(aes(x= Condition, y= d_nonGamer, color= Condition)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_point(position = position_jitter(height = 0, width = 0.2), color= 'black') + 
  ylab("Distance to non-gamers' cluster") + 
  theme(panel.grid.major.x = element_blank(), legend.position = "none")
print(gamer_plot)

# ggsave('Gaming habits.pdf', path= 'Plots', width = 12, height = 8, units = 'cm', useDingbats= FALSE)

rm(gamer, gamer_plot)
```

## Intrinsic Motivation
```{r IM}
# computing subscales for the Intrinsic Motivation
IM <- questionnaire %>%
  mutate(enjoyment =  ((IM01_01 + IM01_05 + IM01_08 + IM01_10 + IM01_14 + IM01_17 + IM01_20)/7),
         competence = ((IM01_04 + IM01_07 + IM01_12 + IM01_16 + IM01_22)/5),
         choice = ((IM01_03 + IM01_11 + IM01_15 + IM01_19 + IM01_21)/5),
         pressure =((IM01_02 + IM01_06 + IM01_09 + IM01_13 + IM01_18)/5)) %>%
  
  select(Observer, Condition, enjoyment, competence, choice, pressure) %>%
  gather(key= "Subscale", value = "Response", -Observer, -Condition)

oneway_permutation_test <- function(subscaleDF){
  test_results <- oneway_test(Response ~ Condition, data= subscaleDF)
  
  
  tibble(Subscale= subscaleDF$Subscale[1], Z= statistic(test_results), pvalue= pvalue(test_results))
}

# Fisher-Pitman permutation test on each subscale, multiple comparions adjustment via Holm's method 
IM %>% 
  group_by(Subscale) %>%
  do(oneway_permutation_test(.)) %>%
  ungroup() %>%
  mutate(`p(adjusted)`= p.adjust(pvalue, method = 'holm')) %>%
  kable(.)


IM_plot <- IM %>% ggplot(aes(x= Condition, y= Response, color= Condition)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_point(position = position_jitter(height = 0, width = 0.2), color= 'black') + 
  facet_grid(. ~ Subscale) + 
  theme(panel.grid.major.x = element_blank(), legend.position = "none")

print(IM_plot)

# ggsave('IM.pdf', IM_plot, path= 'Plots', width = 24, height = 16, units = 'cm', useDingbats= FALSE)

rm('oneway_permutation_test', IM, IM_plot)
```